\documentclass{pgnotes}

\title{Cloud Foundations CT1}

\begin{document}

Cloud Foundations CT1. Answer all questions. Each question = 4 points.

\begin{enumerate}

\item What does the acronym SaaS stand for?
  
\item What does the acronym PaaS stand for?
  
\item What does the acronym IaaS stand for?

\item State the 5 characteristics of cloud computing services. (No explanation necessary)
  
\item List the 3 key components of a Terms-Of-Service agreeemnt (1 word each, no explanation required). 

\item What does SSH stand for?
  
\item What TCP port does SSH run on?

\item SSH provides access to a [ graphical / command-line ] interface. (Select one)
  
\item True or false: Windows Server cannot provide SSH access to PowerShell.
  
\item A user wishes to connect to a server at IP address 193.49.201.18. Her username on her local PC is mary and her account on the remote server is named ``manager''. Write down the SSH command she should type to connect.

\item A user wishes to read json in a file data.json into Powershell data structures in the variable \$data. Write the PowerShell command(s) they should type to achieve this.

\item What protocol can be used to transfer files to/from a remote server using SSH? 

\item S3 is an example of object / block / file storage? (Pick 1 of the 3 options)

\item The AWS CLI has 2 sets of commands for working with S3. One set is ``s3''. What is the other?
  
\item An S3 bucket named accounts holds an object named summary.txt. Write down the S3 URI for this object.

\item Is S3 PaaS or IaaS?

\item What do the letters EC in AWS EC2 stand for?

\item A virtual machine in EC2 is known as? (Answer is 1 word)

\item Is AWS EC2 PaaS or IaaS?

\item An AWS EC2 instance normally uses a virtual hard disk cloned from a starting point.
  What is the starting point called? (Hint: A??)

\item What CLI command will list out the instances running on your account? 
  
\item What component is required to allow a VPC and its components to access the internet?

\item EC2 instances are launched into what VPC component?
  
\item An EC2 instance has a public IP 3.25.102.19 and a private IP in its subnet of 10.0.0.2.
  Which IP address (public or private) will be displayed by the instance OS?

\item Name the EC2 component that controls what traffic is allowed to flow into / out of an EC2 instance.
  
\end{enumerate}

\end{document}